Currently, demonstrations of in-space additive manufacturing are restricted to wire-fed processes. While effective, this constraint limits the versatility of in-space fabrication and, therefore, its potential to support space missions. A major barrier to broader adoption of powder-based methods in space is the challenge of storing and delivering powder in the absence of gravity. This thesis builds on previous investigations into cold spray additive manufacturing (CSAM) to design, test, and evaluate a gravity-independent powder feed system suitable for in-space applications.

A fluidised powder-bed architecture was selected, featuring a pneumatically actuated, gas-permeable piston to compact and push the powder toward an outlet. Drawing inspiration from analogous systems in aerospace propulsion, the design was developed through iterative prototyping and tested under terrestrial conditions. Key challenges addressed included powder jamming the piston as well as an insufficient force on the piston preventing it from effectively compacting the powder at the outlet. These issues were addressed through the use of a flexible dual-cone TPU piston capable of adapting to internal tank imperfections. The system was experimentally tested at inlet pressures between 2.7 and 3.5 bar, yielding mass flow rates between 20 and 60 grams per second, significantly higher than comparable systems.

The system exhibited a reasonably consistent and controllable response to varying inlet pressures, suggesting that the architecture is scalable and responsive to upstream conditions. However, efforts to numerically simulate the design under microgravity conditions were only partially successful due to time and software limitations. While no definitive evidence was found that the system performs equivalently in microgravity, preliminary simulation results suggested continued fluidisation was likely.

Although the design requires further refinement and testing, particularly in reduced-gravity environments, this work demonstrates that a fluidised, piston-driven feed system holds promise for enabling powder-based additive manufacturing beyond Earth. The findings contribute to a growing body of knowledge in gravity-independent powder handling.